### 2024-09-17 base-rates.qmd
20240916 - copyediting

$$
\begin{aligned}
  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R | C) \cdot P(C_i) + P(R | \text{not } C) \cdot [1 - P(C_i)]}
\end{aligned}
$$ {#eq-baseRateUpdated3}

In each case, the formula uses three elements to calculate the probability that the hypothesis is true:

1. the [conditional probability](#sec-conditionalProbability) the likelihood of observing the evidence, $R$, given that the hypothesis, $C$, is true (i.e., $P(R|C)$; [sensitivity](#sec-sensitivity))
1. the [conditional probability](#sec-conditionalProbability) the likelihood of observing the evidence, $R$, given that the hypothesis, $C$, is false (i.e., $P(R | \text{not } C)$; [false positive rate](#sec-falsePositiveRate))
1. the [marginal probability](#sec-baseRate) (base rate) of the event occurring (i.e., the prior probability of the hypothesis, $C$ being true; $P(C_i)$)

@@

Example 3A: Now, consider that a second plan just hit the World Trade Center.
What is the probability that a terror attack occurred given that a second plane hit the World Trade Center?

>>>>

$$
\begin{aligned}
  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R | C) \cdot P(C_i) + P(R | \text{not } C) \cdot [1 - P(C_i)]}
\end{aligned}
$$ {#eq-baseRateAlternative}

In each example, the formula uses three elements to calculate the probability that the hypothesis is true:

1. the [conditional probability](#sec-conditionalProbability) the likelihood of observing the evidence, $R$, given that the hypothesis, $C$, is true (i.e., $P(R|C)$; [sensitivity](#sec-sensitivity))
1. the [conditional probability](#sec-conditionalProbability) the likelihood of observing the evidence, $R$, given that the hypothesis, $C$, is false (i.e., $P(R | \text{not } C)$; [false positive rate](#sec-falsePositiveRate))
1. the [marginal probability](#sec-baseRate) (base rate) of the event occurring (i.e., the prior probability of the hypothesis, $C$, being true; $P(C_i)$)

@@

Example 3B: Now, consider that a second plan just hit the World Trade Center.
What is the probability that a terror attack occurred given that a second plane hit the World Trade Center?
